---
title: "PGoresen_Assignment5"
author: "Patricia Goresen"
date: "10/8/2017"
output: 
  html_document:
    keep_md: true
---

# Question 1
### a.import the file and assign to df
```{r, echo=TRUE}
df <- read.table("~/Desktop/MSDS 6306/Unit 5/Live Assignment/yob2016.txt", header = FALSE, sep = ";", col.names = c("Name", "Gender", "Frequency"))
```
### b. Summary and structure of df
```{r, echo=TRUE}
summary(df)
str(df)
```
### c. Misspelled names
```{r, echo=TRUE}
grep("yyy$", df$Name, value = TRUE)
```
### d. Remove name from df
```{r, echo=TRUE}
y2016 <- subset(df, Name!=grep("yyy$", df$Name, value = TRUE))
```
# Question 2
### a. import the file and assign it to y2015
```{r, echo=TRUE}
y2015 <- read.table("~/Desktop/MSDS 6306/Unit 5/Live Assignment/yob2015.txt", header = FALSE, sep = ",", col.names = c("Name", "Gender", "Frequency"))
```
### b. Display the last ten rows and desribe something you find interesting
```{r, echo=TRUE}
tail(y2015, n=10)
#Something interesting: They all have frequency 5, are males and start with Z
```
### c. Merge y2016 and y2015 by Name column; assign it to final
```{r, echo=TRUE}
final <- merge(y2016, y2015, by = "Name" , all=FALSE)
```
# Question 3 
### a. Create a new column "Total"
```{r, echo=TRUE}
final$total <- final$Frequency.x + final$Frequency.y
sum(final$total)
```
##### 11404228 were given popular names in 2015 and 2016 combined

### b. Sort by total and list top ten names
```{r, echo=TRUE}
final <- final[order(-final$total) , ]
head(final$Name, n=10)
```
### c. Omit boys and list top ten girl's names
```{r, echo=TRUE}
final <- final[with(final, !((Gender.x == "M") | (Gender.y == "M"))), ]
head(final$Name, n=10)
```
### d. Write names and totals to CSV file
```{r, echo=TRUE}
write.csv(final[,c("Name","total")], file="TopGirlNames.csv",row.names=FALSE)
```
# Question 4
### URL for Github: https://github.com/patriciagoresen/SMU-MSDS-Homework/tree/master/Homework%205