---
title: "PGoresen_Assignment4"
author: "Patricia Goresen"
date: "10/2/2017"
output: 
  html_document:
    keep_md: true
---

# Question 1 
###a: install fivethirtyeight package
```{r, echo=TRUE}
##install.packages("fivethirtyeight")
library(fivethirtyeight)
```

###b: Assign eighteenth data set to object df
```{r, echo=TRUE}
data(package = "fivethirtyeight")
df <- college_recent_grads
```

###c: Detailed list of the data sets to write out the URL in a comment to the related news story 
```{r, echo=TRUE}
#vignette("fivethirtyeight", package ="fivethirtyeight")
```
######The URL for this data set is http://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/

###d: Dimensions and column names of df
```{r, echo=TRUE}
str(df)
```
#Question 2 

###a: Column names of data and the number of columns
```{r,echo=TRUE}
colnames(df)
ncol(df)
```
###b: Generating a count of each major category and assigning it to major_count 
```{r,echo=TRUE}
major_count <- table(df$major_category)
major_count
```
###c: Barplot for major_counts
```{r,echo=TRUE}
par(las=2)
barplot(major_count, xlab= "Major name", ylab= "Frequency", main = "Frequencies of Majors Carnevale et al", col= c(1:16), horiz = TRUE)
```

###d: Write fivethirtyeight data to a csv
```{r,echo=TRUE}
write.csv(df, file='College_Grads.csv', row.names = FALSE)
```

#Question 3
### Link to github:
####https://github.com/patriciagoresen/SMU-MSDS-Homework.git

